---
title: "UO Survey of Health and Well-being (SHWell)"
author: "Dani Cosme"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    highlight: tango
    theme: united
    toc: true
    toc_float: 
      collapsed: TRUE
      smooth_scroll: TRUE
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# load libraries
```{r, message=FALSE, warning=FALSE}
if (!require(tidyverse)) {
  install.packages('tidyverse')
}

if (!require(devtools)) {
  install.packages('devtools')
}

if (!require(scorequaltrics)) {
  devtools::install_github('jflournoy/qualtrics')
}

if (!require(knitr)) {
  install.packages('knitr')
}
```

# load aesthetic
```{r}
dc_bw = readRDS('~/dc_bw.Rds')
```

# define variables and paths
```{r, echo = FALSE}
cred_file_location = '~/credentials.yaml.DEFAULT'
cols_to_keep = '(qid|id|Finished|StartDate|EndDate)'
survey_name_filter = '^Aborn'
sid_pattern = ''
identifiableData = c('IPAddress') # exclude when printing duplicates
output_file_dir = '~/Documents/code/dsnlab/FP_scripts/self_report/'
rubric_dir = '/Users/danicosme/Dropbox (PfeiBer Lab)/FreshmanProject/FP_R21/ScoringRubrics/SVC_manuscript/'
```


# access qualtrics data
Filter available surveys based on the filter specified above.  

```{r, results = "hide", }
# load credential file
credentials = scorequaltrics::creds_from_file(cred_file_location)

# filter
surveysAvail = scorequaltrics::get_surveys(credentials)
surveysFiltered = filter(surveysAvail, grepl(survey_name_filter, SurveyName))
```

# clean data
## get survey data

```{r getsurveydata, results = "hide", }
# get data
surveys = scorequaltrics::get_survey_data(surveysFiltered,
                                               credentials, 
                                               pid_col = cols_to_keep) %>%
               filter(!item %in% identifiableData) #filter out identifiable data
```

## clean
### Calculate duration and flag participants that haven't used substances (as they are expected to be faster and missing more items)
```{r}
surveys_dur = surveys %>%
  mutate(StartDate = as.POSIXct(StartDate),
         EndDate = as.POSIXct(EndDate),
         duration = round(difftime(EndDate, StartDate, units = 'mins'), 1)) 

surveys_sub = surveys_dur %>%
  spread(item, value) %>%
  mutate(alcohol = ifelse(ALCOHOL_YN == 1, "yes", "no"),
         cannabis = ifelse(CANNABIS_YN == 1, "yes", "no")) %>%
  gather(item, value, -c("StartDate", "EndDate", "Finished", "duration", "qid", "id", "survey_name", "alcohol", "cannabis"))
```

### Check incomplete surveys

Manually checked, verified whether or not they appear to be good faith responses  

Yes  
* R_OkH6CrL8N62w72F  
* R_1LGyUmjSMqK6gpy  
* R_BR2k23E9qshJWGB  
* R_3vEcKYUXQSSrVdf  
* R_3R3h55GC4JtXcL7  
* R_1DHDEKSfRS1s77A  
* R_2cikmabDllsBn65  
* R_2QRsrhdO1imAanj  
* R_2TzsjF1Iauqatxs

Probably not  
* R_3G33cnkbTnbcTKU --> same response for all items in many survey measures  

No  
* everyone >98% missing  


```{r}
# filter out gated substance use questions for those who have never used those substances
incomplete = surveys_sub %>%
  filter(Finished == '0') %>%
  filter(!(alcohol == "no" & grepl("DDQR|B-YAACQ", item)) & !(cannabis == "no" & grepl("DFAQ_CU|RMPI", item)))

# check the percent missing and view responses manually
incomplete %>%
  mutate(missing = ifelse(value == "", 1, 0)) %>%
  group_by(qid, Finished) %>%
  summarize(n.missing = sum(missing, na.rm = TRUE),
            n.total = n(),
            percent.missing = (n.missing / n.total) * 100,
            duration = mean(duration)) %>%
  arrange(percent.missing)

exclude_incomplete = incomplete %>%
  mutate(missing = ifelse(value == "", 1, 0)) %>%
  group_by(qid, Finished) %>%
  summarize(n.missing = sum(missing, na.rm = TRUE),
            n.total = n(),
            percent.missing = (n.missing / n.total) * 100,
            duration = mean(duration)) %>%
  filter(percent.missing > 98)

surveys_incomplete = surveys_sub %>%
  filter(!qid %in% exclude_incomplete$qid)
```
N incomplete excluded from this step = `r length(unique(exclude_incomplete$qid))`

### Check for very fast or slow responses  

* exclude responses faster than 16 minutes  
* excude responses longer than 48 hours  
```{r}
# all responses
surveys_incomplete %>%
  select(qid, duration, alcohol, cannabis) %>%
  unique() %>%
  mutate(substances = ifelse(alcohol == "no" & cannabis == "no", "0",
                      ifelse(alcohol == "yes" & cannabis == "no", "1",
                      ifelse(alcohol == "no" & cannabis == "yes", "1", "2")))) %>%
  mutate(n = n(),
         mean = mean(duration, na.rm = TRUE),
         lowersd2 = mean - (2 * sd(duration, na.rm = TRUE))) %>%
  ggplot(aes(duration, color = substances)) +
    geom_freqpoly() +
    dc_bw

# responses < 120 mins
plot.data = surveys_incomplete %>%
  select(qid, duration, alcohol, cannabis) %>%
  unique() %>%
  mutate(substances = ifelse(alcohol == "no" & cannabis == "no", "0",
                      ifelse(alcohol == "yes" & cannabis == "no", "1",
                      ifelse(alcohol == "no" & cannabis == "yes", "1", "2")))) %>%
  filter(duration < 100) %>%
  mutate(n = n(),
         mean = mean(duration, na.rm = TRUE),
         lowersd2 = mean - (2 * sd(duration, na.rm = TRUE))) 

plot.data %>%
  ggplot(aes(duration, color = substances)) +
    geom_freqpoly() +
    geom_vline(aes(xintercept = mean)) +
    geom_vline(aes(xintercept = lowersd2), linetype = "dotted") +
    scale_x_continuous(breaks = seq(0, 100, 10)) +
    dc_bw

exclude_fast = surveys_incomplete %>%
  select(qid, duration) %>%
  unique() %>%
  filter(duration < unique(plot.data$lowersd2))

exclude_slow = surveys_incomplete %>%
  select(qid, duration) %>%
  unique() %>%
  filter(duration > (60 * 24 * 2))

surveys_time = surveys_incomplete %>%
  filter(!qid %in% c(exclude_fast$qid, exclude_slow$qid))
```
N fast excluded from this step = `r length(unique(exclude_fast$qid))`

N slow excluded from this step = `r length(unique(exclude_slow$qid))`

### Check for response invariance for numeric responses

* Included as flags of response quality: `sd_questionnaire` and `percent_survey_invariant`   
* `sd_questionnaire` = standard deviation for each questionnaire  
    * If questionnaire was not completed, `sd_questionnaire = NaN`  
    * If questionnaire is a single item, `sd_questionnaire = NA`  
* `percent_survey_invariant` = percent of the survey (across all questionnaires > 1 item completed) that have no variance (SD = 0)
```{r}
invariance = surveys_time %>%
  mutate(questionnaire = gsub("\\d", "", item),
         questionnaire = gsub("_", "", questionnaire),
         questionnaire = gsub("FFMQR", "FFMQ", questionnaire),
         value = as.numeric(value)) %>%
  select(qid, questionnaire, value) %>%
  group_by(qid, questionnaire) %>%
  summarize(sd_questionnaire = sd(value, na.rm = TRUE)) %>%
  group_by(qid) %>%
  mutate(completed_variance = ifelse(!(is.na(sd_questionnaire) | is.nan(sd_questionnaire)), 1, 0),
         n_questionnaires = sum(completed_variance, na.rm = TRUE),
         invariant = ifelse(sd_questionnaire == 0, 1, 0),
         sum_invariant = sum(invariant, na.rm = TRUE),
         percent_survey_invariant = (sum_invariant / n_questionnaires ) * 100) %>%
  select(qid, questionnaire, sd_questionnaire, percent_survey_invariant) %>%
  ungroup()

invariance %>%
  select(qid, percent_survey_invariant) %>%
  unique() %>%
  mutate(mean = mean(percent_survey_invariant, na.rm = TRUE),
         uppersd2 = mean + (2 * sd(percent_survey_invariant, na.rm = TRUE))) %>%
  ggplot(aes(percent_survey_invariant)) +
    geom_histogram(alpha = .5) +
    geom_freqpoly() +
    geom_vline(aes(xintercept = mean)) +
    geom_vline(aes(xintercept = uppersd2), linetype = "dotted") +
    scale_x_continuous(breaks = seq(0, 70, 10)) +
    scale_y_continuous(breaks = seq(0, 120, 10)) +
    dc_bw

surveys_inv = surveys_time %>%
  mutate(questionnaire = gsub("\\d", "", item),
         questionnaire = gsub("_", "", questionnaire),
         questionnaire = gsub("FFMQR", "FFMQ", questionnaire)) %>%
  left_join(., invariance, by = c("qid", "questionnaire"))
```

### Check number of words written in free response questions
* Included as a flag of response quality = `word_count_free_response` 
* `word_count_free_response`= word count across all free response questions (6 total)  
```{r}
word_count = surveys_inv %>%
  filter(grepl("HRB", item)) %>%
  group_by(qid) %>%
  mutate(value = tolower(value),
         word_count = stringr::str_count(value, pattern = "\\w+")) %>%
  select(qid, item, value, word_count) %>%
  group_by(qid) %>%
  summarize(word_count_free_response = sum(word_count, na.rm = TRUE))

word_count %>%
  mutate(mean = mean(word_count_free_response, na.rm = TRUE)) %>%
  ggplot(aes(word_count_free_response)) +
    geom_histogram(alpha = .5) +
    geom_freqpoly() +
    geom_vline(aes(xintercept = mean)) +
    scale_x_continuous(breaks = seq(0, 700, 100)) +
    scale_y_continuous(breaks = seq(0, 120, 10)) +
    dc_bw

surveys_word = surveys_inv %>%
  left_join(., word_count, by = "qid")

surveys_word %>% 
  select(qid, percent_survey_invariant, word_count_free_response) %>% 
  unique() %>% 
  ggplot(aes(word_count_free_response, percent_survey_invariant)) + 
    geom_point(alpha = .3) + 
    geom_smooth() + 
    dc_bw
```

### Check for duplicate responses

* blank responses were recorded before ids were assigned; manually verified there are no duplicates on unique variables (e.g., height/weight) --> treat them as separate entries  
```{r}
surveys_word %>%
  select(-c(questionnaire, sd_questionnaire, percent_survey_invariant)) %>%
  spread(item, value) %>%
  group_by(id) %>%
  summarize(n = n()) %>%
  arrange(desc(n)) %>%
  filter(n > 1) 
```

### Remove duplicate observations and assign new ids
```{r}
surveys_dup = surveys_word %>%
  filter(!qid == "R_rdV3sZiAc89DHyx") %>% #finished response
  filter(!qid == "R_1Qob0KWn8DrLehS") %>% #finished response
  filter(!qid == "R_01IQWZONgCq6NyN") %>% #finished response
  mutate(id = group_indices_(., .dots = "qid"))
```
N duplicates excluded from this step = `r length(unique(surveys_inv$qid)) - length(unique(surveys_dup$qid))`

### Check for non-numeric items using the `get_uncoercibles()` function
```{r}
surveys_dup %>%
  scorequaltrics::get_uncoercibles() %>%
  distinct(item, value) %>%
  arrange(item) %>%
  select(item) %>%
  unique()
  
surveys_dup %>%
  scorequaltrics::get_uncoercibles() %>%
  distinct(item, value) %>%
  arrange(item) %>%
  filter(!grepl("HRB", item))
```

## check manual entry numeric data

* AGE, ALCOHOL_AGE, DDQR5-8, DFAQ_CU_17, DFAQ_CU_22-24, DFAQ_CU_27, DFAQ_CU_30  

(this is still in progress)
```{r}
surveys_dup %>%
  filter(item %in% c("AGE", "ALCOHOL_AGE", "DDQR_5", "DDQR_6", "DDQR_7", "DDQR_8", 
                     "DFAQ_CU_17", "DFAQ_CU_22", "DFAQ_CU_23", "DFAQ_CU_24", "DFAQ_CU_27", "DFAQ_CU_30"))

# check AGE
surveys_dup %>%
  select(qid, item, value) %>%
  filter(item == "AGE") %>%
  mutate(value = as.numeric(value)) %>%
  arrange(desc(value))

# check ALCOHOL_AGE
surveys_dup %>%
  select(qid, item, value) %>%
  filter(item == "ALCOHOL_AGE") %>%
  mutate(value = as.numeric(value)) %>%
  arrange(desc(value))

# check DDQR_5
surveys_dup %>%
  select(qid, item, value) %>%
  filter(item == "DDQR_5") %>%
  mutate(value = as.numeric(value)) %>%
  arrange(desc(value))

# check DDQR_6
surveys_dup %>%
  select(qid, item, value) %>%
  filter(item == "DDQR_6") %>%
  mutate(value = as.numeric(value)) %>%
  arrange(desc(value))

# check DDQR_7
surveys_dup %>%
  select(qid, item, value) %>%
  filter(item == "DDQR_7") %>%
  mutate(value = as.numeric(value)) %>%
  arrange(desc(value))

# check DDQR_8
surveys_dup %>%
  select(qid, item, value) %>%
  filter(item == "DDQR_8") %>%
  mutate(value = as.numeric(value)) %>%
  arrange(desc(value))
```

### Tidy data

* remove messed up surveys: planfulness  
* remove unnecessary items and columns  
* replace missing data with NA  
```{r}
surveys_clean = surveys_dup %>%
  select(-c(survey_name, StartDate, EndDate)) %>%
  filter(!grepl("PLAN_S|ExternalDataReference|intro|consent|debrief|ResponseSet|Status|Name", item)) %>%
  mutate(value = ifelse(value == "", NA, value)) %>%
  rename("finished" = Finished)
```
**Total N before exclusions (including incomplete and duplicate responses) = `r length(unique(surveys$qid))`**

**Total N after exclusions = `r length(unique(surveys_clean$id))`**
